---
title: "BellaBeat_Case_study"
author: "Dhilipkumarkk"
date: "2023-06-13"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction

This case study follows from the capstone project from the Google Data Analytics Course. I will be using the 6 step data analysis process for this case study: Ask, Prepare, Process, Analyze, Share and Act.

##installing libraries

```{r libraries}
library("here")
library("skimr")
library("janitor")
library("dplyr")
library("tidyverse")
```

## creating data frames

```{r data frames}
setwd("C:/Users/kkdhi/Downloads/BellaBeat_Case_Study_2")
daily_activity<-read.csv("dailyActivity_merged.csv")
hourly_calories<-read.csv("hourlyCalories_merged.csv")
hourly_intensities<-read.csv("hourlyIntensities_merged.csv")
hourly_steps<-read.csv("hourlySteps_merged.csv")
sleep_day<-read.csv("sleepDay_merged.csv")
weight_log<-read.csv("weightLogInfo_merged.csv")
```

##merging daily activity dataset with sleep dataset

```{r merging data sets}
daily_activity_with_sleep<-merge(daily_activity,sleep_day,by="Id")
daily_activity_with_wieght<-merge(daily_activity,weight_log,by="Id")
```

##creating summary statistics

```{r summary statistics}
daily_activity %>%  
  select(TotalSteps,
         TotalDistance,
         SedentaryMinutes) %>%
  summary()

sleep_day %>%  
  select(TotalSleepRecords,
         TotalMinutesAsleep,
         TotalTimeInBed) %>%
  summary()

hourly_calories %>%  
  select(ActivityHour, Calories) %>%
  summary()

hourly_intensities %>%  
  select(ActivityHour, TotalIntensity, AverageIntensity) %>%
  summary()

hourly_steps %>%  
  select(ActivityHour, StepTotal) %>%
  summary()

weight_log %>%  
  select(WeightKg, WeightPounds, Fat, BMI, IsManualReport) %>%
  summary()
```

##converting activity date to weekdays

```{r convert date}
daily_activity <- daily_activity %>% mutate( Weekday = weekdays(as.Date(ActivityDate, "%m/%d/%Y")))
daily_activity$weekday1 <- ordered(daily_activity$Weekday, levels=c("Monday","Tuesday","Wednesday","Thursday","Friday","Saturday","Sunday"))
activity_data <- daily_activity %>% 
  group_by(Weekday) %>% 
  summarize(count = n()) 
```

##plotting weekday activity_data

```{r plot activity_data}
ggplot(activity_data, aes(x=Weekday, y=count)) +
  geom_bar(stat="identity",color="black",fill="purple") +
  labs(title="Daily activity usage", x="Days", y="Count")
```

##plotting relationship between steps taken and sedentary minutes

```{r plot realtionship}
ggplot(data=daily_activity, aes(x=TotalSteps, y=SedentaryMinutes)) + 
  geom_point(colour="purple", size=0.5)+geom_smooth(method="lm")+
  labs(title="Relationship between total steps and sedentary minutes", x="Total steps", y="Sedentary minutes")
```

##plotting relationship between total steps taken and calories burnt

```{r plot between step and calories}
ggplot(data=daily_activity, aes(x=TotalSteps, y=Calories))+
  geom_point(color="purple", size=0.5)+ 
  labs(title="Relationship between total steps and calories burnt", x= "Total Steps",
       y = "Calories")+ geom_smooth(method = 'loess', formula= 'y ~ x')
```

##plotting relationship between total minutes asleep and total time in bed

```{r plot}
ggplot(data=sleep_day, aes(x=TotalMinutesAsleep, y=TotalTimeInBed)) +
  geom_point(colour="purple", size=0.5) +
  labs(title="Relationship between time asleep and time in bed",x="Time Asleep",
       y="Time in Bed") +geom_smooth(method="lm")
```

##finding the absolute number of people getting their recommended number of hours of sleep (7-9hours, 420-540min)

```{r finding}
daily_activity_with_sleep %>%
  count(TotalMinutesAsleep < 420)
daily_activity_with_sleep %>%
  count(TotalMinutesAsleep > 540)
```

##changing time format for plotting

```{r format}
hourly_calories$ActivityHour <- parse_date_time(hourly_calories$ActivityHour,
                                                "%m/%d/%Y %I:%M:%S %p")
hourly_calories$date <- as.Date(hourly_calories$ActivityHour, format = "%d/%m/%Y")
hourly_calories$time <- format(as.POSIXct(hourly_calories$ActivityHour), format = "%H:%M:%S")
```

##plotting hourly calories burnt

```{r plot}
ggplot(data=hourly_calories, aes(x=time, y= Calories)) + 
  geom_histogram(stat = "identity", fill="purple") +
  theme(axis.text.x = element_text(angle = 90)) +
  labs(title="Average calories burnt by the hour", x="Time", y="Calories")
```
